---
title: "30-feature-engineering"
output: html_document
---

# 4. Using best practices, write four functions which add engineered features to the dataset, including a description and rationale. Include a test to determine whether the features are working as intended. 

##In order to understand some pitch measurements better, let's provide some context by finding the standard score for release_speed, release_spin_rate, pfx_x, & pfx_z for each pitch, as thrown by either a right handed pitcher or a left handed pitcher.  This will let us see how each pitch relates to every other based on these measures:
```{r standardized measures}
scale1 <- function(x) round((x - mean(x))/ sd(x), digits = 3)

std_by_hand <- function(df, vars) {
  df %>% 
    group_by(p_throws) %>% 
    mutate_at(vars, list(std = scale1)) %>% 
    ungroup()%>% 
    select(contains("std"), everything())
}

fastballs %>% 
  std_by_hand(c("release_speed", "release_spin_rate", "pfx_x", "pfx_z")) %>% 
  assert(within_bounds(-4, 4, pfx_z_std, pfx_x_std, release_spin_rate_std))
```


##A basic way to measure the effectivness of a pitch is to observe whether the batter hit the ball or missed the ball with his swing.  Let's pull this data from the column that stores this and make a new column, "result".  Note that pitches that are missed are still swings, so the total number of swings is the sum of "swing" and "miss".
```{r results}
fastballs %>% 
  mutate(result = case_when(description %in% c("foul", "hit_into_play", "hit_into_play_no_out", "hit_into_play_score") ~ "swing",  description %in% c("foul_tip", "swinging_strike") ~ "miss"))
```


##
```{r avg by pitcher}
mean1 <- function(x) round((mean(x)), digits = 3)

avg_by_pitcher <- function(
  
  
  
  
  
  
```

 
 
##Can we find each pitcher's max for each of our four measurements, release_speed, release_spin_rate, pfx_x, & pfx_z? Group by "pitcher":
```{r max by pitcher}
max1 <- function(x) round((max(x)), digits = 3)

max_by_pitcher <- function(
  
  
  
  
  
  
  
  
```
